---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# codeconsensus

<!-- badges: start -->
<!-- badges: end -->

The goal of `codeconsensus` is to assist with electronic health record (EHR) code curation from mutiple EHR dictionaries.


## Installation
You can install the development version of codeconsensus from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("nicksunderland/codeconsensus")
```

## Concepts
At the heart of `codeconsensus` is the specification of a `concept`. A `concept` is the clinical entity which we are trying to describe in terms of diagnostic codes. Concepts are defined in a `.yaml` file with the following structure, for example Heart Failure. The concept `.yaml` files are located in project-specific directories within this package at `/inst/concepts/...`.

The primary field of concern is that `regexes` field - this needs to be adjusted with appropriate search terms. The regexes can be for all dictionaries (add under the `all` field) or specific to a particular dictionary/terminology. 

```{yaml}
id: esc_heart_failure
name: Heart Failure
definition: |
  Heart failure is not a single pathological diagnosis, but a clinical syndrome consisting of
  cardinal symptoms (e.g. breathlessness, ankle . swelling, and fatigue) that may be accompanied
  by signs (e.g. elevated . jugular venous pressure, pulmonary crackles, and peripheral oedema).
  It is due to a structural and/or functional abnormality of the heart that results in elevated
  intracardiac pressures and/or inadequate cardiac output at rest and/or during exercise.
reference: https://doi.org/10.1093/eurheartj/ehab368
domain: Disorder
terminology:
- SNOMED
- ICD11
- ICD10
regexes:
  SNOMED: '84114007'
  ICD10:
  - ~
  all:
  - (?i)(heart|cardiac|ventricular) failure
  - (?i)cardiac insufficiency
  - ' HF '
  ICD11:
  - ~
include: esc_heart_failure
exclude: ~
```


## EHR dictionaries
The concept files are used to query the appropriate EHR dictionaries, the dictionaries used are those specified under the `terminology` field. The EHR dictionary files can be large and so are not included in this package. The dictionaries include: 

* ICD 10  
* ICD 11  
* ICD 9  
* SNOMED  
* CPT4  
* OPCS  

Currently the dictionary directory is located at: 

```{yaml}
'/path-to-UCL-OneDrive/.../IHI.CompMed - Documents/PHENOTYPING/ehr_dictionaries'
```


## Create code lists
To create a concept code list you can use the function `make_concept()`. This will output the code list in the 3 requested formats:

* Flat `.tsv` file (list option)  
* Excel file called `concepts.xlsx`, where concepts are added as tabs across multiple function calls (excel option)  
* Code tree `.RDS` file (nested lists), suitable for displaying in Shiny applications (tree option)  

```{r ehr_dictionaries, eval = FALSE}
library(codeconsensus)

concept_yml <- system.file("concepts", "nih_cardiomyopathy", "nih_heart_failure.yaml", package="codeconsensus")

make_concept(concept_yml   = concept_yml, 
             output_dir    = "~/Desktop/codeconsensus_output", 
             ehr_dict_root = Sys.getenv("CODECONSESUS_DICT_DIR"),
             format        = c("list", "excel", "tree"))
```

## Code conensus
The workflow after generating an initial code list includes: 

* manual comparison with published code lists  
* expert manual review of the codes, voting, then concensus meeting  
* amendment of the regexes to capture additional codes  
* rerunning of the `make_concept()` function to generate the final code list

